Logistic Regression on Baxter colorectal cancer dataset

### 2018-08-30 LogisticRegression_BaxterData.Rmd
epochs=100 will run the whole thing x100 times

#### Start by importing modules that will be necessary 
```{python,tidy = TRUE, cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
############## IMPORT MODULES ######################
from sklearn.linear_model import LogisticRegression
from sklearn import linear_model
from sklearn.model_selection import train_test_split
import numpy as np # linear algebra
import pandas as pd # data processing, CSV file I/O (e.g. pd.read_csv)
from sympy import *
import matplotlib.pyplot as plt
from scipy import interp
from sklearn.metrics import roc_curve, auc
from sklearn.model_selection import RepeatedStratifiedKFold
from sklearn.model_selection import GridSearchCV
```

@@

We have 6920 OTUs that are the features we put in.
We have 292 samples with labels:
- 120 Cancer colon
- 172 Normal colon

@@

## Drop if NA elements
y.dropna()
x.dropna()

## Split dataset to 80% training 20% test sets.

x_train, x_test, y_train, y_test = train_test_split(x, y, test_size=0.2,shuffle=True)

#### Now let's define a L2 regularized logistic regression model
- We tune our hyper-parameter(regualarization strength) here with 80-20 splits and 1000 repeats. 
- Once we have our "C" value tuned, we define our best model.
```{python,tidy = TRUE,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
######################### Logistic Regression ##############################

@@

## Define the n-folds for hyper-parameter optimization on training set.
cv = RepeatedStratifiedKFold(n_splits=5, n_repeats=1000, random_state=200889)
## We will try these regularization strength coefficients to optimize our model
C = {"C": [0.0000001, 0.000001, 0.00001, 0.0001]}
## Define the best model:
clf = GridSearchCV(logreg, C, cv=cv, verbose=0, scoring='roc_auc')
best_model = clf.fit(x_train, y_train)
print('Best C:', best_model.best_estimator_.get_params()['C'])
print('Best model:', best_model.best_estimator_)
## The best model we pick here will be used for predicting test set.
best_model = best_model.best_estimator_

#### Plot ROC curve for cross validation of training set and then test set 
1. Plot mean ROC curve for cross-validation with n_splits=5 (80-20 stratified) and n_repeats=1000 to evaluate the variation of prediction in our training set.
2. Plot ROC curve for never before seen test set.

```{python,tidy = TRUE,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
## Now let's plot some ROC curves
tprs = []
aucs = []
mean_fpr = np.linspace(0, 1, 100)

tprs_test = []
aucs_test = []
mean_fpr_test = np.linspace(0, 1, 100)

@@

## Plot mean ROC curve for cross-validation with n_splits=5 and n_repeats=1000 to evaluate the variation of prediction in our training set.
for train, test in cv.split(X,Y):
    probas_ = best_model.fit(X[train], Y[train]).predict_proba(X[test])
    fpr, tpr, thresholds = roc_curve(Y[test], probas_[:, 1])
    tprs.append(interp(mean_fpr, fpr, tpr))
    tprs[-1][0] = 0.0
    roc_auc = auc(fpr, tpr)
    aucs.append(roc_auc)

## Predict and plot ROC curve for never before seen test set.
probas_ = best_model.predict_proba(x_test)
# Compute ROC curve and area the curve
fpr_test, tpr_test, thresholds_test = roc_curve(y_test, probas_[:, 1])
tprs_test.append(interp(mean_fpr_test, fpr_test, tpr_test))
tprs_test[-1][0] = 0.0
roc_auc_test = auc(fpr_test, tpr_test)
aucs_test.append(roc_auc_test)

plt.plot([0, 1], [0, 1], linestyle='--', color='green', label='Luck', alpha=.8)
mean_tpr_test = np.mean(tprs_test, axis=0)
mean_tpr_test[-1] = 1.0
mean_auc_test = auc(mean_fpr_test, mean_tpr_test)
std_auc_test = np.std(aucs_test)
plt.plot(mean_fpr_test, mean_tpr_test, color='r', label=r'Never-before-seen test set ROC (AUC = %0.2f $\pm$ %0.2f)' % (mean_auc_test, std_auc_test), lw=2, alpha=.8)
std_tpr_test = np.std(tprs_test, axis=0)
tprs_upper_test = np.minimum(mean_tpr_test + std_tpr_test, 1)
tprs_lower_test = np.maximum(mean_tpr_test - std_tpr_test, 0)
plt.fill_between(mean_fpr_test, tprs_lower_test, tprs_upper_test, color='tomato', alpha=.2, label=r'$\pm$ 1 std. dev.')
mean_tpr = np.mean(tprs, axis=0)
mean_tpr[-1] = 1.0
mean_auc = auc(mean_fpr, mean_tpr)
std_auc = np.std(aucs)
plt.plot(mean_fpr, mean_tpr, color='b', label=r'Mean cross-val ROC (AUC = %0.2f $\pm$ %0.2f)' % (mean_auc, std_auc), lw=2, alpha=.8)
std_tpr = np.std(tprs, axis=0)
tprs_upper = np.minimum(mean_tpr + std_tpr, 1)
tprs_lower = np.maximum(mean_tpr - std_tpr, 0)
plt.fill_between(mean_fpr, tprs_lower, tprs_upper, color='dodgerblue', alpha=.2, label=r'$\pm$ 1 std. dev.')
plt.xlim([-0.05, 1.05])
plt.ylim([-0.05, 1.05])
plt.xlabel('False Positive Rate')
plt.ylabel('True Positive Rate')
plt.title('L2 Logistic Regression ROC\n')
plt.legend(loc="lower right", fontsize=8)
plt.show()
#Logit_plot.savefig('results/figures/Logit_Baxter.png', dpi=1000)

@@

>>>>

#### Start by importing modules that will be necessary 
```{python,tidy = TRUE, cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
############## IMPORT MODULES ######################
from sklearn.linear_model import LogisticRegression
from sklearn import linear_model
from sklearn.model_selection import train_test_split
import numpy as np # linear algebra
import pandas as pd # data processing, CSV file I/O (e.g. pd.read_csv)
from sympy import *
import matplotlib.pyplot as plt
from scipy import interp
from sklearn.metrics import roc_curve, auc
from sklearn.model_selection import RepeatedStratifiedKFold
from sklearn.model_selection import GridSearchCV
from sklearn.externals import joblib
```

@@

We have 6920 OTUs that are the features we put in.
We have 292 samples with labels:

  - 120 Cancer colon
  - 172 Normal colon

@@

## Drop if NA elements
y.dropna()
x.dropna()
```

#### Now let's define a L2 regularized logistic regression model
1. We will select hyper-parameters and an optimal model during training with 80% of the data and test that best model on the remanining %20 x100 times.
  - The only hyper-parameter in Logit here will be the strength of the L2-regularization
2. For each epoch, the chosen best model will be tested on the %20 test set that was not seen before during trainig. We will report mean AUC values +/- sd for the testing/predicting.
3. For each epoch, we will also be splitting the 80% training data 5-fold again, another 100 times and cross validate (inner loop). We will report mean AUC values +/- sd for each cross-validation.
- 
```{python,tidy = TRUE,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
## Generate empty lists to fill with AUC values for test-set
tprs_test = []
aucs_test = []
mean_fpr_test = np.linspace(0, 1, 100)

######################### Logistic Regression ##############################

epochs= 100
for epoch in range(epochs):
    ## Split dataset to 80% training 20% test sets.
    x_train, x_test, y_train, y_test = train_test_split(x, y, test_size=0.2,shuffle=True)

@@

    ## Define the n-folds for hyper-parameter optimization on training set.
    cv = RepeatedStratifiedKFold(n_splits=5, n_repeats=100, random_state=200889)
    ## We will try these regularization strength coefficients to optimize our model
    C = {"C": [0.0000001, 0.000001, 0.00001, 0.0001]}
    ## Define the best model:
    clf = GridSearchCV(logreg, C, cv=cv, verbose=0, scoring='roc_auc')
    best_model = clf.fit(x_train, y_train)
    print('Best C:', best_model.best_estimator_.get_params()['C'])
    print('Best model:', best_model.best_estimator_)
    ## The best model we pick here will be used for predicting test set.
    best_model = best_model.best_estimator_

    ## Generate empty lists to fill with AUC values for train-set cv
    tprs = []
    aucs = []
    mean_fpr = np.linspace(0, 1, 100)

@@

    ## Predict never-before-seen test-set
    probas_ = best_model.predict_proba(x_test)
    # Compute ROC curve and area the curve
    fpr_test, tpr_test, thresholds_test = roc_curve(y_test, probas_[:, 1])
    tprs_test.append(interp(mean_fpr_test, fpr_test, tpr_test))
    tprs_test[-1][0] = 0.0
    roc_auc_test = auc(fpr_test, tpr_test)
    aucs_test.append(roc_auc_test)
    print("Test", roc_auc_test)

    ## Plot mean ROC curve for cross-validation with n_splits=5 and n_repeats=100 to evaluate the variation of prediction in our training set.
    for train, test in cv.split(X,Y):
        probas_ = best_model.fit(X[train], Y[train]).predict_proba(X[test])
        fpr, tpr, thresholds = roc_curve(Y[test], probas_[:, 1])
        tprs.append(interp(mean_fpr, fpr, tpr))
        tprs[-1][0] = 0.0
        roc_auc = auc(fpr, tpr)
        aucs.append(roc_auc)
        print("Train", roc_auc)

@@

#### Plot ROC curve for cross validation of training set and test set

```{python,tidy = TRUE,comment='[Output]', cache=FALSE, message=FALSE, warning=FALSE, engine.path = '/Library/Frameworks/Python.framework/Versions/3.6/bin/python3'}
plt.plot([0, 1], [0, 1], linestyle='--', color='green', label='Luck', alpha=.8)
mean_tpr_test = np.mean(tprs_test, axis=0)
mean_tpr_test[-1] = 1.0
mean_auc_test = auc(mean_fpr_test, mean_tpr_test)
std_auc_test = np.std(aucs_test)
plt.plot(mean_fpr_test, mean_tpr_test, color='r', label=r'Never-before-seen test set ROC (AUC = %0.2f $\pm$ %0.2f)' % (mean_auc_test, std_auc_test), lw=2, alpha=.8)
std_tpr_test = np.std(tprs_test, axis=0)
tprs_upper_test = np.minimum(mean_tpr_test + std_tpr_test, 1)
tprs_lower_test = np.maximum(mean_tpr_test - std_tpr_test, 0)
plt.fill_between(mean_fpr_test, tprs_lower_test, tprs_upper_test, color='tomato', alpha=.2, label=r'$\pm$ 1 std. dev.')
mean_tpr = np.mean(tprs, axis=0)
mean_tpr[-1] = 1.0
mean_auc = auc(mean_fpr, mean_tpr)
std_auc = np.std(aucs)
plt.plot(mean_fpr, mean_tpr, color='b', label=r'Mean cross-val ROC (AUC = %0.2f $\pm$ %0.2f)' % (mean_auc, std_auc), lw=2, alpha=.8)
std_tpr = np.std(tprs, axis=0)
tprs_upper = np.minimum(mean_tpr + std_tpr, 1)
tprs_lower = np.maximum(mean_tpr - std_tpr, 0)
plt.fill_between(mean_fpr, tprs_lower, tprs_upper, color='dodgerblue', alpha=.2, label=r'$\pm$ 1 std. dev.')
plt.xlim([-0.05, 1.05])
plt.ylim([-0.05, 1.05])
plt.xlabel('False Positive Rate')
plt.ylabel('True Positive Rate')
plt.title('L2 Logistic Regression ROC\n')
plt.legend(loc="lower right", fontsize=8)
plt.show()
#Logit_plot.savefig('results/figures/Logit_Baxter.png', dpi=1000)
